---
title: "US Craft Beer and Brewery Analysis"
author: "Fred Poon, Laura Lazarescou"
date: "10/21/2019"
output:
  powerpoint_presentation: default
  slidy_presentation: default
  ioslides_presentation: default
subtitle: Presented to Carlos Brito, CEO and Felipe Dutra, CFO
---

```{r setup, include=FALSE}

# knitr options
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warn = FALSE)
knitr::opts_chunk$set(message = FALSE)
# set root directory to project root

# put it in the directory that contains the ENTIRE PROJECT
# commented out for this project:  knitr::opts_knit$set(root.dir = "..")

# load libraries

library(tidyverse)
library(usmap)

library(stringr)
library(class)
library(caret)
library(e1071)
```
## Beer Data: Quality and Consistency
* Approximately half of the beer data does not include IBU values (1005/2410) values. 
* IBU is not a true predictor of taste.
* Buyers do not base decisions on IBU and consumers don't care. 
* ABV was not reported for 64 beers but they also lacked IBU so we omitted them.

## Breweries Across the United States

```{r, fig.width=10, fig.height=8}

# get beers data
beers = read.csv("D:/OneDrive - Southern Methodist University/6306 doing data science/bivinsadler-msds/MDS-6306-Doing-Data-Science-Fall-2019/Unit 8 and 9 Case Study 1/Beers.csv", header = TRUE)

# get breweries data
breweries = read.csv("D:/OneDrive - Southern Methodist University/6306 doing data science/bivinsadler-msds/MDS-6306-Doing-Data-Science-Fall-2019/Unit 8 and 9 Case Study 1/Breweries.csv", header = TRUE)
# breweries = as_tibble(breweries)

# get consumption data
consumption = read.csv("D:/OneDrive - Southern Methodist University/6306 doing data science/bivinsadler-msds/MDS-6306-Doing-Data-Science-Fall-2019/Unit 8 and 9 Case Study 1/Consumption.csv",header = TRUE)
names(consumption)[names(consumption) == "ï..StateCd"] <- "StateCd"

# breweries %>% group_by(State) %>% tally() %>% arrange(desc(n))
breweries %>% 
  count(State, sort=TRUE) %>% 
  ggplot(aes(x = reorder(State, -n), y = n)) + 
    geom_bar(stat = "identity", fill = "#aa66cd") +
    geom_text(aes(label = n), vjust = - 1) +
    ggtitle("Breweries Across the United States") +
    ylab("Number of Breweries") +
    xlab("State") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=90, vjust = 1))

# group by state
# breweries %>% ggplot(aes(State)) + 
#    geom_bar(fill= "#875D2D") + 
#    geom_text(stat="count", aes(label=..count..), hjust=-1, size=2.75) + 
#    ggtitle("Breweries Across the United States") +
#    ylab("Number of Breweries") +
#    coord_flip() + 
#    theme(axis.text=element_text(size = 12),axis.title=element_text(size=14, face="bold"),title = element_text(size=16, #face="bold"))

```

## Top 10 States for Brewing

```{r}
#  Question 1 continued
#  Showing top 10 / bottom 10
# top 10
breweries %>% group_by(State) %>% tally() %>% arrange(desc(n)) %>%
  head(n = 10) %>%
  ggplot(aes(reorder(State,n),n)) +
  geom_bar(stat = "identity", fill = "#66cdaa") +
  ylab("Number of Breweries") +
  xlab("State") + 
  ggtitle("Most Breweries per State") +
  coord_flip() +
  theme(axis.text=element_text(size = 12),axis.title=element_text(size=14, face="bold"),title = element_text(size=16, face="bold"))
```

## States with the Least Number of Breweries
```{r}
# bottom 10
breweries %>% group_by(State) %>% tally() %>% arrange(desc(n)) %>%
  tail(n = 10) %>%
  ggplot(aes(reorder(State,-n),n)) +
  geom_bar(stat = "identity", fill = "#66bdcd") +
  ylab("Number of Breweries") +
  xlab("State") + 
  ggtitle("Least Number of Breweries")+
  coord_flip() + 
  theme(axis.text=element_text(size = 12),axis.title=element_text(size=14, face="bold"),title = element_text(size=16, face="bold"))


## Question 2
#Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  #(RMD only, this does not need to be included in the presentation or the deck.)
# outer join to merge the data set

bb = merge(beers,breweries, by.x = "Brewery_id", by.y = "Brew_ID", all=TRUE)

# head(bb)
# tail(bb)


## Question 3
#Address the missing values in each column. - answered in Slide 2 above

# separating data set to omit na's
bb2 = na.omit(bb)
# summary(bb2)

```

## Median Alcohol (ABV) and Bitterness (IBU) for Each State

```{r,fig.width = 15}
# Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

bb2 %>% 
  group_by(State) %>% 
  summarize(medianABV = median(ABV) * 100, medianIBU = median(IBU)) %>%
  gather('medianABV','medianIBU', key = "category", value = "value") %>%
  ggplot(aes(x = reorder(State,-value), y = value)) +
  geom_bar(stat = "identity", aes(fill = category)) +
  geom_text(aes(label = value), angle = 90, hjust =-1) + 
  ylab("Median ABV and Median IBU") + xlab("State") + ggtitle("Median ABV and IBU for Each State") +
  theme(axis.text=element_text(size = 12),axis.title=element_text(size=14, face="bold"),title = element_text(size=16, face="bold"),legend.text = element_text(size = 11), axis.text.x=element_text(angle=90))

# trying with a 100% scale - is it easier to see a correlation?
# bb2 %>% 
#  group_by(State) %>% 
#  summarize(medianABV = median(ABV) * 100, medianIBU = median(IBU)) %>%
#  gather('medianABV','medianIBU', key = "category", value = "value") %>%
#   ggplot(aes(x = reorder(State,-value), y = value, fill = category)) +
#   geom_bar(position = "fill", stat = "identity") +
#   scale_y_continuous(labels = scales::percent_format()) +
#   ylab("Percent Median of ABV to Median of IBU") + xlab("State") + ggtitle("Median of ABV to IBU Comparison") +
#   theme(axis.text=element_text(size = 12),axis.title=element_text(size=14, face="bold"),title = element_text(size=16, 
# face="bold"),legend.text = element_text(size = 11))

```

## Median Alcohol (ABV) by State
```{r}
mapdata <- bb2 %>%
  group_by(State) %>% 
  summarize(medianABV = median(ABV) * 100, medianIBU = median(IBU))
names(mapdata)[names(mapdata) == "State"] <- "state"
names(mapdata)[names(mapdata) == "medianABV"] <- "abv"
names(mapdata)[names(mapdata) == "medianIBU"] <- "ibu"
mapdata$state = as.character(mapdata$state)
mapdata$state = str_trim(mapdata$state)

plot_usmap(data = mapdata, values = "abv", color = "#000000") + 
  scale_fill_continuous(
    low = "#bdcd66", high = "#4b5228", name="Median ABV by State", label = scales::comma
  ) + labs(title = "Median ABV by State") + theme(title=element_text(size=8), legend.position = "right")

```

## Median Bitterness (IBV) by State

```{r}

plot_usmap(data = mapdata, values = "ibu", color = "#000000") + 
  scale_fill_continuous(
    low = "#cdaa66", high = "#524428", name="Median IBU by State", label = scales::comma
  ) + labs(title = "Median IBU by State") + theme(title=element_text(size=8), legend.position = "right")

```

## State with Beer having Highest Alcohol by Volume (ABV)

``` {r}
# Which state has the maximum alcoholic (ABV) [Alcohol by Volume] beer? Which state has the most bitter (IBU) [International 
# Bitterness Units] beer?
maxABV <- bb %>% 
  select(State, ABV) %>% 
  arrange(desc(ABV))

head(maxABV, n = 1)
```
## State with the Most Bitter Beer

```{r}

maxIBU <- bb %>% 
  select(State, IBU) %>% 
  arrange(desc(IBU))

head(maxIBU,n = 1)

```

## Top Beer Consuming States

```{r}

withconsumption <- merge(mapdata,consumption, by.x = "state", by.y = "StateCd", all=TRUE)
withconsumption %>% arrange(desc(Barrels)) %>% head(n=15) %>%
  ggplot(aes(x = abv, y = ibu)) + 
  geom_point(aes(color = state, size = Barrels), alpha = 0.9) +
  scale_size(range = c(0.5, 25)) +
  geom_text(label="California - Most Consumption", x=4, y=62, alpha = .75) + 
  geom_text(label="Maine - Most IBU and ABV", x=4, y=60, alpha=.75) + ggtitle("Top Beer Consuming States")

```

## ABV Distribution

```{r, fig.width=8}
#summary(bb$ABV)

bb2$ABVPerc = bb2$ABV * 100
bb2 %>% 
  ggplot(aes(ABVPerc)) + 
  geom_histogram(color = "#000000", fill = "#cd6689", bins= 51) + 
  ggtitle("ABV Distribution") + 
  ylab("Frequency") + 
  xlab("ABV %") + 
  theme(axis.text=element_text(size = 12),axis.title=element_text(size=14, face="bold"),title = element_text(size=16, face="bold"))

```

## ABV & IBU by Style of Beer

```{r, fig.width=7, fig.height=6}

# create a new category for general styles of beers
bb2$NewStyle = 
  ifelse(grepl("lager", bb2$Style, ignore.case = TRUE), "Lager",
    ifelse(grepl("india pale ale", bb2$Style, ignore.case = TRUE),"IPA",
      ifelse(grepl("ipa", bb2$Style, ignore.case = TRUE),"IPA", 
        ifelse(grepl("cider", bb2$Style, ignore.case = TRUE), "Cider",
          ifelse(grepl("ale",bb2$Style, ignore.case = TRUE), "Ale",
            ifelse(grepl("stout", bb2$Style, ignore.case = TRUE), "Porter/Stout",
              ifelse(grepl("pilsner", bb2$Style, ignore.case = TRUE), "Pilsner",
                ifelse(grepl("pilsener", bb2$Style, ignore.case = TRUE), "Pilsner",
                  ifelse(grepl("porter", bb2$Style, ignore.case = TRUE), "Porter/Stout",
                    ifelse(grepl("fruit", bb2$Style, ignore.case = TRUE), "Fruit/Herb",
                      ifelse(grepl("herbed", bb2$Style, ignore.case = TRUE), "Fruit/Herb",
                        ifelse(grepl("hefeweizen", bb2$Style, ignore.case = TRUE), "Wheat",
                          ifelse(grepl("wheat", bb2$Style, ignore.case = TRUE), "Wheat","Other")))))))))))))

# bb2 %>% filter(NewStyle == "Other")
# bb2 %>% group_by(NewStyle) %>% count() %>% arrange(desc(n))



# View ABV as a percentage
bb2$ABVPerc = bb2$ABV * 100

bb2 %>%
  ggplot(aes(x = ABVPerc, y = IBU, color=NewStyle)) +
  geom_point() +
  ggtitle("ABV & IBU by Style of Beer") +
  theme(axis.text=element_text(size = 12),axis.title=element_text(size=14, face="bold"),title = element_text(size=16, face="bold"),legend.text = element_text(size = 11))+xlab("% ABV")

# plot(bb)

```

## KNN

```{r}

AleIPA <- bb2 %>% filter(NewStyle == "IPA" | NewStyle == "Ale") %>% select(NewStyle,ABV,IBU)
#--view(AleIPA)

set.seed(6)
perc=.75
trainIndices = sample(1:dim(AleIPA)[1],round(perc * dim(AleIPA)[1]))
train = AleIPA[trainIndices,]
test = AleIPA[-trainIndices,]
iterations=30

accs = data.frame(accuracy = numeric(iterations), k = numeric(iterations))

for(i in 1:iterations)
{
  classifications = knn(train[,c(2,3)],test[,c(2,3)],train$NewStyle, prob = TRUE, k = i)
  table(test$NewStyle,classifications)
  matrix = confusionMatrix(table(test$NewStyle,classifications))
  accs$accuracy[i] = matrix$overall[1]
  accs$k[i] = i
}


AleIPA %>%
  ggplot(aes(x = ABV, y = IBU, color=NewStyle)) +
  geom_point() +
  ggtitle("ABV & IBU by Style of Beer") +
  theme(axis.text=element_text(size = 12),axis.title=element_text(size=14, face="bold"),title = element_text(size=16, face="bold"),legend.text = element_text(size = 11))+xlab("% ABV")


matrix

```